---
title: "Weirdness of People and Methods of Weird"
author: "Dr. Fletcher"
date: "April 16, 2019"
output:
  html_document: default
  pdf_document: default
---

# PSYCHOMETRIC MODELING: "The weirdness of people and methods of weird"

## STRUCTURAL EQUATION MODELING (SEM)
```{r}
library(tidyverse)
library(lavaan)
library(devtools)
library(BAS)
setwd("C:/WORK")

testdat <- read.csv("C:/WORK/unconference project.csv")
names(testdat) <- tolower(names(testdat))
testdat.subset <- testdat %>% select(starts_with("TT"))

# Filtering for complete cases
selected_cols <- testdat %>% select(starts_with("TT")) %>% filter(complete.cases(.))

data_pca <- prcomp(selected_cols, center = TRUE, scale = TRUE)

summary(data_pca)

data_pca$rotation

plot(data_pca)

summary(selected_cols)

fit3 <- factanal(selected_cols, 5, rotation="varimax")
print(fit3, digits=2, cutoff=.3, sort=TRUE)

#remove those with low factor loadings and those with non-simple structure
trimmed.fields <- subset(testdat.subset, select = -c(tt.eng.direct.mail,tt.eng.broadcast.cable.tv,tt.dms.recreational.shoppers,tt.conv.etail.only,tt.adch.online.video,tt.dms.novelty.seekers,tt.adch.online.streaming.tv,tt.dms.in.the.moment.shoppers,tt.eng.mobile.sms.mms,tt.eng.radio,tt.eng.traditional.newspaper,tt.conv.discount.supercenters,tt.conv.online.deal.voucher,tt.adch.broadcast.cable.tv,tt.adch.internet.radio,tt.adch.satellite.radio,tt.conv.ebid.sites,tt.conv.etailer,tt.conv.online.bid.marketplace,tt.conv.online.discount,tt.adch.mobile.video,tt.tp.it.s.all.in.the.name,tt.tp.never.show.up.empty.handed,tt.adch.online.display,tt.conv.brick.mortar))

#run CFA on reduced item set
fit <- factanal(na.omit(trimmed.fields), 4, scores = c("regression"), rotation="varimax")
print(fit, digits=2, cutoff=.3, sort=TRUE)

#My lavaan model
#NO DUPLICATES BETWEEN FACTORS
model <- '
  # measurement model
factor1=~tt.tp.buy.american+tt.dms.brand.loyalists+tt.adch.direct.mail+tt.dms.trendsetters+tt.tp.look.at.me.now+tt.tp.on.the.road.again+
tt.eng.digital.newspaper+tt.dms.mainstream.adopters+tt.dms.savvy.researchers+tt.adch.mobile.display+tt.eng.streaming.tv+
tt.eng.digital.video+tt.eng.digital.display

factor2=~tt.dms.deal.seekers+tt.dms.mainstream.adopters+tt.conv.online.mid.high+tt.adch.email.engagement+tt.conv.specialty.dept.store+
tt.conv.wholesale+tt.eng.digital.newspaper+tt.conv.mid.high.end.store+tt.conv.specialty.or.boutique

factor3=~tt.conv.specialty.dept.store+tt.conv.wholesale+tt.conv.online.mid.high+tt.tp.stop.and.smell.the.roses+tt.dms.deal.seekers+
tt.dms.mainstream.adopters+tt.tp.go.with.the.flow+tt.tp.show.me.the.money+tt.tp.no.time.like.the.present+
tt.tp.penny.saved.a.penny.earned+tt.tp.work.hard..play.hard

factor4=~tt.conv.specialty.dept.store+tt.eng.digital.newspaper+tt.conv.wholesale+tt.dms.quality.matters+tt.dms.trendsetters+
tt.dms.organic.and.natural

factor1 ~~ factor1
factor2 ~~ factor2 
factor3 ~~ factor3
factor4 ~~ factor4
'

#comment out bc can't run on full dataset
fitsem <- sem(model, na.omit(trimmed.fields))
summary(fitsem, standardized=TRUE)
```

## LINEAR MIXED METHODS (LMM)
```{r setup, include=FALSE}
library(haven)
depression <- read_sav("depression.sav")
View(depression)
library(haven)
Neighborhood <- read_sav("Neighborhood.sav")
View(Neighborhood)
library(haven)
School <- read_sav("School.sav")
View(School)
```

```{r}
library(lme4)
library(ggplot2)
library(lmerTest)
library(sjPlot)
library(sjstats)
library(ggeffects)
library(snakecase)

#Histogram
ggplot(Neighborhood, aes(x=distress)) +
  geom_histogram(bins = 50)

#Linear Mixed-Effects Models, neigbhorhood and distress
base <- lmer(distress ~ 1 + (1|cluster), data = Neighborhood, REML = FALSE)
summary(base)
plot_model(base, type = "diag")
plot_model(base, type = "re")

#Intraclass correlation coefficient for LMM model
ICC <- icc(base)

#Linear Mixed-Effects Models, life event, distress, and support
randint <- lmer(distress ~ levent + support + (1|cluster), REML = F, data = Neighborhood)
summary(randint)
plot_model(randint, type = "diag")
plot_model(randint, type = "re")
plot_model(randint, type = "est")
plot_model(randint, type = "pred")

#ANOVA Test #1
anova1 <- anova(base, randint)
summary(anova1)

#Random Intersect #1
ri <- ggpredict(randint, c("levent", "cluster [sample=10]"), type = "re")
summary(ri)
plot(ri)

randint2 <- lmer(distress ~ levent + support + (1 + levent|cluster), REML = F, data = Neighborhood)
summary(randint2)
plot(randint2)
plot_model(randint2, type = "diag")
plot_model(randint2, type = "re")
plot_model(randint2, type ="est")
plot_model(randint2, type = "pred")

#ANOVA TEST #2
anova2 <- anova(randint, randint2)
summary(anova2)

#Random Intersect #2
ri2 <- ggpredict(randint2, c("support", "cluster [sample=10]"), type = "re")
summary(ri2)
plot(ri2)

randint3 <- lmer(distress ~ levent + support + zndisord + (1|cluster), REML = F, data = Neighborhood)
summary(randint3)
plot(randint3)
plot_model(randint3, type = "diag")
plot_model(randint3, type = "re")
plot_model(randint3, type = "est")
plot_model(randint3, type = "pred")

#Random Intersect #3
ri3 <- ggpredict(randint3, c("zndisord", "cluster [sample=10]"), type = "re")
summary(ri3)
plot(ri3)

#ANOVA TEST #3
anova3 <- anova(randint2, randint3)

```

## APPENDICES

### STRUCTURAL EQUATION MODELING (SEM) & LINEAR MIX METHODS (LMM)

Tutorials on lavaan 
<http://lavaan.ugent.be/tutorial/tutorial.pdf>; 
<http://lavaan.ugent.be/tutorial/syntax1.html>;
<https://www.jstatsoft.org/article/view/v048i02/v48i02.pdf>

An introduction on Structural Equation Modeling
<http://joophox.net/publist/semfamre.pdf>

Structural Equation Modeling Examples
<http://sachaepskamp.com/semPlot/examples>

Data-sets for SEM Modeling
UCIrvine data sets
<https://archive.ics.uci.edu/ml/index.php>

Google Data Sets
<https://www.google.com/publicdata/directory>

KD Nuggets Data List
<https://www.kdnuggets.com/datasets/index.html>

Student Performance Data with time components
<https://rstudio-pubs-static.s3.amazonaws.com/108835_65a73467d96f4c79a5f808f5b8833922.html>

U.S. Gov't Open data
<http://www.data.gov/open-gov/>

Yelp dataset challenge
<https://www.yelp.com/dataset>

Oh Canada! Open Gov't data
<http://open.canada.ca/en/open-data>

Nonlinear Dimensionality Reduction by John A Lee & Michael Verleysen
<https://amzn.to/2GfOg7Q>

Modern Multivariate Statistical Techniques: Regression, Classification, and Manifold Learning by Alan Julian Izenman
<https://amzn.to/2KEdSAr>

Two-way Interactions
<https://strengejacke.github.io/sjPlot/articles/plot_interactions.html#two-way-interactions>

Best-Practice Recommendations for Estimating Cross-Level Interaction Effects Using Multilevel Modeling
<http://www.hermanaguinis.com/JOMcrosslevel.pdf>

Testing and Interpreting Interaction Effects in Multilevel Models
<https://www.ncaase.com/docs/Stevens__Schulte_AERA_2016_final.pdf>

Two-Level Hierarchical Linear Models Using SAS, Stata, HLM, R, SPSS, and Mplus
<https://stat.utexas.edu/images/SSC/Site/hlm_comparison-1.pdf>