---
title: "Linear Mixed Methods: School Dataset"
output: html_notebook
---

## Load libraries

```{r}
library(haven) # To read SPSS files
library(lme4)
library(lmerTest)
library(sjPlot)
library(sjstats)
library(ggplot2)
```


## Read data

```{r}
school <- read_spss("School.sav")
school <- as.data.frame(school)
summary(school)
```

Response variable:

* langpost = language score at end of year

School-level effects:

* MeanIQverb = mean verbal IQ of the school
* MeanSES = mean SES of school

Individual effects (scores centered w/in their respective schools):

* CSES = group mean centered (with school as reference) SES score
* CVerbIQ = group mean centered (with school as reference) verbal IQ score

Note: To get a student's individual score SES, for example, you need to add MeanSES + CSES

Ideal data setup:

* Level 1 data: each individual student's data is group means centered against others within their school
* Level 2 data: each individual school is grand means centered against all other schools in the data


## Exploratory

```{r}
ggplot(school, aes(x=langpost)) +
  geom_histogram(bins=50)

school$schoolnr <- factor(school$schoolnr) # Convert schoolnr to factor

ggplot(school, aes(y=langpost, group=schoolnr)) +
  geom_boxplot()
```


## Model 1: Baseline model with random intercept effects for school

```{r}
base_model <- lmer(langpost ~ 1 + (1 | schoolnr), REML=FALSE, data=school)
summary(base_model)
```

Does school affiliation help account for some of the variance in the model?  Check the ICC.  It's > 0, so we've got a good case for a mixed effects model.

```{r}
icc(base_model)
```

Some of our schools look like they have very low intercept estimates--and the confidence intervals don't cross zero, so these are likely low-performing schools.

```{r}
plot_model(base_model, type="diag")
plot_model(base_model, type="re")
```


## Model 2: Model with random intercept for school AND individual fixed effects for student attributes

```{r}
model2 <- lmer(langpost ~ CSES + CVerbIQ + (1 | schoolnr), REML=FALSE, data=school)
summary(model2)
```


```{r}
anova(base_model, model2)
```


## Model 3: Model with random intercept for school AND individual fixed effects for student attributes AND random slope effects for CVerbIQ by school

```{r}
model3 <- lmer(langpost ~ CSES + CVerbIQ + (1 + CVerbIQ | schoolnr), REML=FALSE, data=school)
summary(model3)
```

Plot to help interpret the random slopes that vary by school:

Plot ideas: https://strengejacke.github.io/sjPlot/articles/plot_interactions.html#two-way-interactions

```{r}
library(ggeffects)

plot <- ggpredict(model3, c("CVerbIQ", "schoolnr [1, 2, 10, 12, 15, 18, 47, 103]"), type = "re")

plot(plot)

plot_model(model3, type="re")
```


```{r}
ranef(model3)
```





